The Rockerverse: Images, Packages, and Applications for Containerization with R

### 2020-01-20 manuscript.Rmd
minor changes to Dodona use case

  - name: Craig Willis
    # https://github.com/craig-willis
      - 501 E. Daniel St.
      - \orcid{0000-0002-6148-7196}
    email: willis8@illinois.edu

@@

**Whole Tale** (https://wholetale.org) is a National Science Foundation (NSF) funded project\footnote{NSF Award OAC-1541450} developing a scalable, open-source, multi-user platform for reproducible research \citep{brinckman2019, chard2019a}.
A central goal of the platform is to enable researchers to easily create and publish executable research objects\footnote{In Whole Tale a `tale` is a research object that contains metadata, data (by copy or reference), code, narrative, documentation, provenance, and information about the computational environment to support computational reproducibility} associated with published research \citep{chard2019b}.
For R users, Whole Tale attempts to combine the strengths of the Rocker Project's curated Docker images with the Jupyter Project's `repo2docker`\footnote{See \href{https://repo2docker.readthedocs.io/}{https://repo2docker.readthedocs.io/}} component to simplify the customization of R-based environments for researchers with limited experience with either Docker or Git.
Using Whole Tale, researchers can create and publish Rocker-based reproducible research objects to a growing number of repositories including DataONE member nodes, Zenodo and soon Dataverse.
Additionally, Whole Tale supports automatic data citation and is working on capabilities for image preservation and provenance capture to improve the transparency of publish computational research artifacts \citep{mecum2018, mcphillips2019}.

@@

@@

## Publication Reproducibility Audits

A recommendation of the recent National Academies' report on _Reproducibility and Replicability in Science_ is that journals "consider ways to ensure computational reproducibility for publications that make claims based on computations" \citep{NASEM2019}.
In fields such as Political Science and Economics, journals are increasingly adopting policies requiring authors to publish the code and data required to reproduce computational findings reported in published manuscripts, subject to independent verification \citep{Jacoby2017,Vilhuber2019,Alvarez2018,Christian2018,Eubank2016,King1995}.
Problems with the computational environment, installation and availability of software dependencies are common. R is gaining popularity in these communities (in a sample of 105 replication packages published by the _American Journal of Political Science_ (AJPS) over 65% use R). 
Projects such as NSF Whole Tale use the Rocker Project community images with the goal of improving the reproducibility of published research artifacts and simplifying the publication and verification process for both authors and journals by reducing errors and time spent specifying the environment.

@@

The ordering of authors following DN as first author is alphabetical. DN conceived the article idea, \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised all sections.
DB wrote the section on \pkg{outsider}.
GD contributed the CARD.com use case.
DE wrote the introduction and the section about Containerization and Rocker.
RC \& EH contributed to the section on interfaces for Docker in R.
DC contributed content on Gigantum.
ME contributed to the section on processing and deployment to cloud services.
CF wrote paragraphs about \pkg{r-online}, \pkg{dockerfiler}, \pkg{r-ci} and \pkg{r-db}.
BM wrote the section on research compendia and made the project Binder-ready.
HN \& JN co-wrote the section on enterprise production environments (\pkg{googleComputeEngineR}, Google Cloud Run, Kubernetes).
KR wrote the section about \pkg{wholepunch}.
NR wrote paragraphs about shared work environments and GPUs.
LS \& NT wrote the section on Bioconductor.
CW wrote the sections on Whole Tale and contributed the publication reproducibility audit use case.
NX wrote the section on \pkg{liftr}.
All authors approved the final version. This articles was collaboratively written at \href{https://github.com/nuest/rockerverse-paper/}{https://github.com/nuest/rockerverse-paper/}.
The \href{https://github.com/nuest/rockerverse-paper/graphs/contributors}{contributors page}, \href{https://github.com/nuest/rockerverse-paper/commits/master}{commit history}, and \href{https://github.com/nuest/rockerverse-paper/issues/}{discussion issues} provide a detailed view on the respective contributions.

@@

DN is supported by the project Opening Reproducible Research ([o2r](https://www.uni-muenster.de/forschungaz/project/12343)) funded by the German Research Foundation (DFG) under project number [PE&nbsp;1632/17-1](https://gepris.dfg.de/gepris/projekt/415851837).
The funders had no role in data collection and analysis, decision to publish, or preparation of the manuscript.
CW is supported by the Whole Tale projects (https://wholetale.org) funded by the US National Science Foundation (NSF) under award [OAC-1541450](https://www.nsf.gov/awardsearch/showAward?AWD_ID=1541450).

>>>>

@@

@@

Another way this idea of a common environment and the partitioning that Docker provides is used in teaching is in the **[Dodona](https://dodona.ugent.be)** project.
Dodona is a web platform developed at Ghent University and is used to teach students basic programming skills.
Dodona uses Docker containers to test submissions by students.
This means that both the code testing the students' submissions and the submission itself are executed in a predictable environment, avoiding compatibility issues between the wide variety of configurations used by students.
The containerization is also used to shield the Dodona servers from bad (or even malicious) code: memory, time and I/O limits are used to make sure students can't overload the system.
The web application communicates with containers by sending configuration information as a JSON document over standard input. 
Every Dodona Docker image shares a `main.sh` file that passes through this information to the actual testing framework, while setting up some error handling.
The testing process in the Docker containers sends back the test results by writing a JSON document to its standard output channel.
In June 2019, R support was added to Dodona using an image derived from the `rocker/r-base` image that sets up the `runner` user and and `main.sh` file expected by Dodona\footnote{\href{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}}. 
It also installs the packages required for the testing framework and the exercises so that this doesn't have to happen every time a student's submission is evaluated.
The actual testing of R exercises is done using a custom framework loosely based on \CRANpkg{testthat}.
During the development of the testing framework it was found that the \CRANpkg{testthat} framework did not provide enough information to its reporter system to send back all the fields required by Dodona to render its feedback.
Right now, multiple statistics courses are developing exercises to automate the feedback for their lab classes.

@@

@@

The ordering of authors following DN as first author is alphabetical. DN conceived the article idea, \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised all sections.
DB wrote the section on \pkg{outsider}.
GD contributed the CARD.com use case.
DE wrote the introduction and the section about Containerization and Rocker.
RC \& EH contributed to the section on interfaces for Docker in R.
DC contributed content on Gigantum.
ME contributed to the section on processing and deployment to cloud services.
CF wrote paragraphs about \pkg{r-online}, \pkg{dockerfiler}, \pkg{r-ci} and \pkg{r-db}.
BM wrote the section on research compendia and made the project Binder-ready.
HN \& JN co-wrote the section on enterprise production environments (\pkg{googleComputeEngineR}, Google Cloud Run, Kubernetes).
KR wrote the section about \pkg{wholepunch}.
NR wrote paragraphs about shared work environments and GPUs.
LS \& NT wrote the section on Bioconductor.
NX wrote the section on \pkg{liftr}.
All authors approved the final version. This articles was collaboratively written at \href{https://github.com/nuest/rockerverse-paper/}{https://github.com/nuest/rockerverse-paper/}.
The \href{https://github.com/nuest/rockerverse-paper/graphs/contributors}{contributors page}, \href{https://github.com/nuest/rockerverse-paper/commits/master}{commit history}, and \href{https://github.com/nuest/rockerverse-paper/issues/}{discussion issues} provide a detailed view on the respective contributions.

@@

DN is supported by the project Opening Reproducible Research ([o2r](https://www.uni-muenster.de/forschungaz/project/12343)) funded by the German Research Foundation (DFG) under project number [PE&nbsp;1632/17-1](https://gepris.dfg.de/gepris/projekt/415851837).
The funders had no role in data collection and analysis, decision to publish, or preparation of the manuscript.
